Supplementary Material

### 2023-10-24 ms_preprint/ms_supp_mat.Rmd
ms adjustments after LB RS and MS reviews pre-submission

---
title: "Supplementary Material"
subtitle: "<b>Seabird assemblages are linked to the major western boundary current off eastern Australia</b>"
# author: "Author"
date: "`r Sys.Date()`"
output:
  pdf_document:
    keep_tex: true
    # fig_caption: false
bibliography: ../references.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

@@

The code is archived in a repository **(Zenodo repo)**, and you can find a detailed walk-through in it.

@@

![**Figure S1.** Number of occurrences (nOCC), frequency of occurrence (FO) and numeric frequency (NF) of seabirds recorded off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021. The dashed line represents the number of occurrence thresholds (n = 6) each taxon had to match for its inclusion in the seasonal models (see *Methods* in the main text). Species are ordered from the largest to the lowest total number of occurrences.](../results/FigS1_spp-nOCC-FO-NF-seasons.png)

@@

![**Figure S2.** Species richness and the total number of seabirds counted off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021, by grid/season, after data aggregation (see *Methods* in the main text). Dots represent the grid centroids.](../results/FigS2_map-effort-spp-richness-and-total-birds-seasons.png)

@@

![**Figure S3.1.** Multifit plot for Region of Common Profiles (RCP) for each seasonal presence-absence model, applied to seabirds off eastern Australia. The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data. For each number of groups, we ran 100 models with random starting values to avoid getting stuck in an incorrect 'optima' (see *Methods* in the main text). The resulting plot also shows how many groups were 'empty' (colour scale) with 'five or less' or 'only one' sites assigned to an RCP, i.e. the model was fit with, say, 5 groups, but 3 of them had 'five or less' or 'only one' sites (grids) allocated to an RCP.](../results/FigS3_1_multifit-Bernoulli.png)

@@

![**Figure S3.2.** Multifit plot for Region of Common Profiles (RCP) for each seasonal abundance (count) model, applied to seabirds off eastern Australia. The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data. For each number of groups, we ran 100 models with random starting values to avoid getting stuck in an incorrect 'optima' (see *Methods* in the main text). The resulting plot also shows how many groups were 'empty' (colour scale) with 'five or less' or 'only one' sites assigned to an RCP, i.e. the model was fit with, say, 5 groups, but 3 of them had 'five or less' or 'only one' sites (grids) allocated to an RCP.](../results/FigS3_2_multifit-NegBin.png)

@@

![**Figure S4.1.** Best model residuals for each seasonal presence-absence Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_1_best-model-residuals-Bernoulli.png)

@@

![**Figure S4.2.** Best model residuals for each seasonal abundance (count) Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_2_best-model-residuals-NegBin.png)

@@

![**Figure S5.1.** Partial plots for the retained covariates in the best seasonal models based on presence-absence data. The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value. Refer to Table 1 in the main text for the environmental data acronyms.](../results/FigS6_1_partial-plots-Bernoulli.png)

@@

![**Figure S5.2.** Partial plots for the retained covariates in the best seasonal models based on abundance data. The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value. Refer to Table 1 in the main text for the environmental data acronyms.](../results/FigS6_2_partial-plots-NegBin.png)

@@

![**Figure S6.1.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from presence-absence models. The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_1_prob-pred-Bernoulli.png)

@@

![**Figure S6.2.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from abundance (count) models. The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_2_prob-pred-NegBin.png)

@@

![**Figure S7.1.** Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal presence-absence model. Values are the average and confidence intervals of probability of occurrence for each species, based on 1000 Bayesian bootstraps. ](../results/FigS7_1_spp-profiles-Bernoulli.png)

@@

![**Figure S7.2.** Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal abundance (count) model. Values are the average and confidence intervals of predicted mean abundance for each species, based on 1000 Bayesian bootstraps. Values were log10-transformed to accommodate the high variation between species.](../results/FigS7_2_spp-profiles-NegBin.png)

@@

![**Figure S8.** Diversity curve (alpha diversity) and sample coverage for each assemblage (Region of Common Profile; RCP) from each presence-absence seasonal model.](../results/FigS8_iNEXT.png)

>>>>

---
title: "Supplementary Material"
subtitle: "Seabird assemblages are linked to the major western boundary current off eastern Australia"
# author: "Author"
date: "`r Sys.Date()`"
output:
  pdf_document:
    keep_tex: true
    # fig_caption: false
bibliography: ../references.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

@@

The code is archived in a repository **(Zenodo)**, where you can find a detailed walk-through.

@@

![**FIGURE S1.** Number of occurrences (nOCC), frequency of occurrence (FO) and numeric frequency (NF) of seabirds recorded off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021. The dashed line represents the number of occurrence thresholds (n = 6) each taxon had to match for its inclusion in the seasonal models (see *Methods* in the main text). Species are ordered from the largest to the lowest total number of occurrences.](../results/FigS1_spp-nOCC-FO-NF-seasons.png)

@@

![**FIGURE S2.** Species richness and the total number of seabirds counted off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021, by grid/season, after data aggregation (see *Methods* in the main text). Dots represent the grid centroids.](../results/FigS2_map-effort-spp-richness-and-total-birds-seasons.png)

@@

![**FIGURE S3.1.** Multifit plot for Region of Common Profiles (RCP) for each seasonal presence-absence model, applied to seabirds off eastern Australia. The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data. For each number of groups, we ran 100 models with random starting values to avoid getting stuck in an incorrect 'optima' (see *Methods* in the main text). The resulting plot also shows how many groups were 'empty' (colour scale) with 'five or less' or 'only one' sites assigned to an RCP, i.e. the model was fit with, say, 5 groups, but 3 of them had 'five or less' or 'only one' sites (grids) allocated to an RCP.](../results/FigS3_1_multifit-Bernoulli.png)

@@

![**FIGURE S3.2.** Multifit plot for Region of Common Profiles (RCP) for each seasonal abundance (count) model, applied to seabirds off eastern Australia. The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data. For each number of groups, we ran 100 models with random starting values to avoid getting stuck in an incorrect 'optima' (see *Methods* in the main text). The resulting plot also shows how many groups were 'empty' (colour scale) with 'five or less' or 'only one' sites assigned to an RCP, i.e. the model was fit with, say, 5 groups, but 3 of them had 'five or less' or 'only one' sites (grids) allocated to an RCP.](../results/FigS3_2_multifit-NegBin.png)

@@

![**FIGURE S4.1.** Best model residuals for each seasonal presence-absence Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_1_best-model-residuals-Bernoulli.png)

@@

![**FIGURE S4.2.** Best model residuals for each seasonal abundance (count) Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_2_best-model-residuals-NegBin.png)

@@

![**FIGURE S5.1.** Partial plots for the retained covariates in the best seasonal models based on presence-absence data. The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value. Refer to Table 1 in the main text for the environmental data acronyms.](../results/FigS6_1_partial-plots-Bernoulli.png)

@@

![**FIGURE S5.2.** Partial plots for the retained covariates in the best seasonal models based on abundance data. The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value. Refer to Table 1 in the main text for the environmental data acronyms.](../results/FigS6_2_partial-plots-NegBin.png)

@@

![**FIGURE S6.1.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from presence-absence models. The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_1_prob-pred-Bernoulli.png)

@@

![**FIGURE S6.2.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from abundance (count) models. The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_2_prob-pred-NegBin.png)

@@

![**FIGURE S7.1.** Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal presence-absence model. Values are the average and confidence intervals of probability of occurrence for each species, based on 1000 Bayesian bootstraps. ](../results/FigS7_1_spp-profiles-Bernoulli.png)

@@

![**FIGURE S7.2.** Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal abundance (count) model. Values are the average and confidence intervals of predicted mean abundance for each species, based on 1000 Bayesian bootstraps. Values were log10-transformed to accommodate the high variation between species.](../results/FigS7_2_spp-profiles-NegBin.png)

@@

![**FIGURE S8.** Diversity curve (alpha diversity) and sample coverage for each assemblage (Region of Common Profile; RCP) from each presence-absence seasonal model.](../results/FigS8_iNEXT.png)

### 2023-10-24 ms_preprint/ms_supp_mat.tex
line numbering and a few corrections after EJWs review, and changed to correct MS title in the supp mat
ms adjustments after LB RS and MS reviews pre-submission

\title{Supplementary Material}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Probabilistic bioregion models suggest seabird assemblages are
linked to the major boundary current off eastern Australia}
\author{}
\date{\vspace{-2.5em}2023-10-17}

@@

The code is archived in a repository \textbf{(Zenodo repo)}, and you can
find a detailed walk-through in it.

@@

\begin{figure}
\centering
\includegraphics{../results/FigS1_spp-nOCC-FO-NF-seasons.png}
\caption{\textbf{Figure S1.} Number of occurrences (nOCC), frequency of
occurrence (FO) and numeric frequency (NF) of seabirds recorded off
eastern Australia during Australasian Seabird Group's ship-based surveys
between 2016--2021. The dashed line represents the number of occurrence
thresholds (n = 6) each taxon had to match for its inclusion in the
seasonal models (see \emph{Methods} in the main text). Species are
ordered from the largest to the lowest total number of occurrences.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS2_map-effort-spp-richness-and-total-birds-seasons.png}
\caption{\textbf{Figure S2.} Species richness and the total number of
seabirds counted off eastern Australia during Australasian Seabird
Group's ship-based surveys between 2016--2021, by grid/season, after
data aggregation (see \emph{Methods} in the main text). Dots represent
the grid centroids.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS3_1_multifit-Bernoulli.png}
\caption{\textbf{Figure S3.1.} Multifit plot for Region of Common
Profiles (RCP) for each seasonal presence-absence model, applied to
seabirds off eastern Australia. The number of groups with the lowest BIC
value indicates the best number of groups (assemblages) that describes
the data. For each number of groups, we ran 100 models with random
starting values to avoid getting stuck in an incorrect `optima' (see
\emph{Methods} in the main text). The resulting plot also shows how many
groups were `empty' (colour scale) with `five or less' or `only one'
sites assigned to an RCP, i.e.~the model was fit with, say, 5 groups,
but 3 of them had `five or less' or `only one' sites (grids) allocated
to an RCP.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS3_2_multifit-NegBin.png}
\caption{\textbf{Figure S3.2.} Multifit plot for Region of Common
Profiles (RCP) for each seasonal abundance (count) model, applied to
seabirds off eastern Australia. The number of groups with the lowest BIC
value indicates the best number of groups (assemblages) that describes
the data. For each number of groups, we ran 100 models with random
starting values to avoid getting stuck in an incorrect `optima' (see
\emph{Methods} in the main text). The resulting plot also shows how many
groups were `empty' (colour scale) with `five or less' or `only one'
sites assigned to an RCP, i.e.~the model was fit with, say, 5 groups,
but 3 of them had `five or less' or `only one' sites (grids) allocated
to an RCP.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS4_1_best-model-residuals-Bernoulli.png}
\caption{\textbf{Figure S4.1.} Best model residuals for each seasonal
presence-absence Region of Common Profile model, applied to seabirds off
eastern Australia.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS4_2_best-model-residuals-NegBin.png}
\caption{\textbf{Figure S4.2.} Best model residuals for each seasonal
abundance (count) Region of Common Profile model, applied to seabirds
off eastern Australia.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS6_1_partial-plots-Bernoulli.png}
\caption{\textbf{Figure S5.1.} Partial plots for the retained covariates
in the best seasonal models based on presence-absence data. The plot
shows the probability of belonging to a Region of Common Profiles (RCP)
against the environmental value. Refer to Table 1 in the main text for
the environmental data acronyms.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS6_2_partial-plots-NegBin.png}
\caption{\textbf{Figure S5.2.} Partial plots for the retained covariates
in the best seasonal models based on abundance data. The plot shows the
probability of belonging to a Region of Common Profiles (RCP) against
the environmental value. Refer to Table 1 in the main text for the
environmental data acronyms.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS5_1_prob-pred-Bernoulli.png}
\caption{\textbf{Figure S6.1.} Predicted probability membership of for
each seabird assemblage (Region of Common Profiles; RCP) and grid, off
eastern Australia, from presence-absence models. The central column,
`mean', corresponds to the point prediction and Bayesian bootstraped,
lower and upper confidence intervals (CI), on its sides.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS5_2_prob-pred-NegBin.png}
\caption{\textbf{Figure S6.2.} Predicted probability membership of for
each seabird assemblage (Region of Common Profiles; RCP) and grid, off
eastern Australia, from abundance (count) models. The central column,
`mean', corresponds to the point prediction and Bayesian bootstraped,
lower and upper confidence intervals (CI), on its sides.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS7_1_spp-profiles-Bernoulli.png}
\caption{\textbf{Figure S7.1.} Species profiles for each assemblage
(Region of Common Profiles; RCP) for each seasonal presence-absence
model. Values are the average and confidence intervals of probability of
occurrence for each species, based on 1000 Bayesian bootstraps.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS7_2_spp-profiles-NegBin.png}
\caption{\textbf{Figure S7.2.} Species profiles for each assemblage
(Region of Common Profiles; RCP) for each seasonal abundance (count)
model. Values are the average and confidence intervals of predicted mean
abundance for each species, based on 1000 Bayesian bootstraps. Values
were log10-transformed to accommodate the high variation between
species.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS8_iNEXT.png}
\caption{\textbf{Figure S8.} Diversity curve (alpha diversity) and
sample coverage for each assemblage (Region of Common Profile; RCP) from
each presence-absence seasonal model.}
\end{figure}

>>>>

\title{Supplementary Material}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Seabird assemblages are linked to the major western boundary
current off eastern Australia}
\author{}
\date{\vspace{-2.5em}2023-10-24}

@@

The code is archived in a repository \textbf{(Zenodo)}, where you can
find a detailed walk-through.

@@

\begin{figure}
\centering
\includegraphics{../results/FigS1_spp-nOCC-FO-NF-seasons.png}
\caption{\textbf{FIGURE S1.} Number of occurrences (nOCC), frequency of
occurrence (FO) and numeric frequency (NF) of seabirds recorded off
eastern Australia during Australasian Seabird Group's ship-based surveys
between 2016--2021. The dashed line represents the number of occurrence
thresholds (n = 6) each taxon had to match for its inclusion in the
seasonal models (see \emph{Methods} in the main text). Species are
ordered from the largest to the lowest total number of occurrences.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS2_map-effort-spp-richness-and-total-birds-seasons.png}
\caption{\textbf{FIGURE S2.} Species richness and the total number of
seabirds counted off eastern Australia during Australasian Seabird
Group's ship-based surveys between 2016--2021, by grid/season, after
data aggregation (see \emph{Methods} in the main text). Dots represent
the grid centroids.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS3_1_multifit-Bernoulli.png}
\caption{\textbf{FIGURE S3.1.} Multifit plot for Region of Common
Profiles (RCP) for each seasonal presence-absence model, applied to
seabirds off eastern Australia. The number of groups with the lowest BIC
value indicates the best number of groups (assemblages) that describes
the data. For each number of groups, we ran 100 models with random
starting values to avoid getting stuck in an incorrect `optima' (see
\emph{Methods} in the main text). The resulting plot also shows how many
groups were `empty' (colour scale) with `five or less' or `only one'
sites assigned to an RCP, i.e.~the model was fit with, say, 5 groups,
but 3 of them had `five or less' or `only one' sites (grids) allocated
to an RCP.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS3_2_multifit-NegBin.png}
\caption{\textbf{FIGURE S3.2.} Multifit plot for Region of Common
Profiles (RCP) for each seasonal abundance (count) model, applied to
seabirds off eastern Australia. The number of groups with the lowest BIC
value indicates the best number of groups (assemblages) that describes
the data. For each number of groups, we ran 100 models with random
starting values to avoid getting stuck in an incorrect `optima' (see
\emph{Methods} in the main text). The resulting plot also shows how many
groups were `empty' (colour scale) with `five or less' or `only one'
sites assigned to an RCP, i.e.~the model was fit with, say, 5 groups,
but 3 of them had `five or less' or `only one' sites (grids) allocated
to an RCP.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS4_1_best-model-residuals-Bernoulli.png}
\caption{\textbf{FIGURE S4.1.} Best model residuals for each seasonal
presence-absence Region of Common Profile model, applied to seabirds off
eastern Australia.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS4_2_best-model-residuals-NegBin.png}
\caption{\textbf{FIGURE S4.2.} Best model residuals for each seasonal
abundance (count) Region of Common Profile model, applied to seabirds
off eastern Australia.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS6_1_partial-plots-Bernoulli.png}
\caption{\textbf{FIGURE S5.1.} Partial plots for the retained covariates
in the best seasonal models based on presence-absence data. The plot
shows the probability of belonging to a Region of Common Profiles (RCP)
against the environmental value. Refer to Table 1 in the main text for
the environmental data acronyms.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS6_2_partial-plots-NegBin.png}
\caption{\textbf{FIGURE S5.2.} Partial plots for the retained covariates
in the best seasonal models based on abundance data. The plot shows the
probability of belonging to a Region of Common Profiles (RCP) against
the environmental value. Refer to Table 1 in the main text for the
environmental data acronyms.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS5_1_prob-pred-Bernoulli.png}
\caption{\textbf{FIGURE S6.1.} Predicted probability membership of for
each seabird assemblage (Region of Common Profiles; RCP) and grid, off
eastern Australia, from presence-absence models. The central column,
`mean', corresponds to the point prediction and Bayesian bootstraped,
lower and upper confidence intervals (CI), on its sides.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS5_2_prob-pred-NegBin.png}
\caption{\textbf{FIGURE S6.2.} Predicted probability membership of for
each seabird assemblage (Region of Common Profiles; RCP) and grid, off
eastern Australia, from abundance (count) models. The central column,
`mean', corresponds to the point prediction and Bayesian bootstraped,
lower and upper confidence intervals (CI), on its sides.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS7_1_spp-profiles-Bernoulli.png}
\caption{\textbf{FIGURE S7.1.} Species profiles for each assemblage
(Region of Common Profiles; RCP) for each seasonal presence-absence
model. Values are the average and confidence intervals of probability of
occurrence for each species, based on 1000 Bayesian bootstraps.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS7_2_spp-profiles-NegBin.png}
\caption{\textbf{FIGURE S7.2.} Species profiles for each assemblage
(Region of Common Profiles; RCP) for each seasonal abundance (count)
model. Values are the average and confidence intervals of predicted mean
abundance for each species, based on 1000 Bayesian bootstraps. Values
were log10-transformed to accommodate the high variation between
species.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../results/FigS8_iNEXT.png}
\caption{\textbf{FIGURE S8.} Diversity curve (alpha diversity) and
sample coverage for each assemblage (Region of Common Profile; RCP) from
each presence-absence seasonal model.}
\end{figure}
